Rubin Observatory Pointing Algorithm and Simulations

### 2025-01-31 ms.tex
Updates from Overleaf

xxx-Cite Eric's paper \citep{Bellm2019} and LCOGT \citep{Lampoudi2015} on integer programming.

@@

@@

\subsection{Kinematic Model}

Run through the kinematic model. Which motions are modeled, cable wrap tracked, etc.

Optional to include a finite jerk for each component using motion profiles from \citet{Mu2009}.

including scheduled and unscheduled downtime---xxx details of each.

@@

@@

For the majority of observations we use a modified Markov Decision Process (MDP), similar to the one developed in \citet{Naghib2019}.

@@

@@

\subsection{}

@@

\software{Astropy \citep{astropy2013, astropy2018}, \\
          Healpy and HEALpix\footnote{\url{http://healpix.sourceforge.net}} \citep{healpix2005, Zonca2019},
          Numpy \citep{harris2020array}, \\
          Matplotlib \citep{Hunter:2007},
          Scipy \citep{2020SciPy-NMeth},
          Pandas \citep{reback2022}, \\
          Scikit-learn \citep{scikit-learn},
          OpenOrb \citep{Granvik2009}, 
          GNU Parallel \citep{Tange2023}, \\
          rubin\_scheduler \citep{rubin_scheduler2024},
          rubin\_sim \citep{rubin_sim2024}}

>>>>

xxx-Cite Eric's paper for ZTF \citep{Bellm2019} and LCOGT paper \citep{Lampoudi2015} on integer programming.

@@

Throughout this paper we adopt the LSST nomenclature of a "visit" being potentially mulitple exposures at a single location while each of those exposures is refered to as a "snap". Rubin processing typically combines snaps from a visit before running things like transient detection.

@@

To simulate the operations of Rubin Observatory, we have constructed a model of the telescope and site which can mimic the telemetry stream to the scheduling software and record the relevant metadata for completed observations. In addition, the model observatory can simulate the relevant motions of the telescope.

\subsection{Kinematic Model}
The kinematic model computes the time in takes for various motions including
\begin{itemize}
    \item{Azimuthal motions of the dome}
    \item{Altitude motion of the dome mounted wind screen}
    \item{Altitude and azimuthal motions of the telescope}
    \item{Rotational motion of the camera}
    \item{Filter change time}
    \item{telescope settle times}
    \item{filter change}
    \item{shutter motion during a visit}
    \item{open and closed optics loop delays}
\end{itemize}

Times for altitude, azimuth, and rotational motions are calculated using a constant acceleration value and maximum velocity (configurable for each component). A finite value for the jerk can be included based on \citet{Mu2009}, but is turned off by default as it has minimal impact on the total motion times.--xxx is this actually true? Looks like we do have jerk on for telescope motions.

xxx--Cable wrap is tracked for the telescope (which can go xxx) and the camera which has limits of xxx.

xxx-figures with starting/ending values, colored by time, for each component.

\subsection{Downtimes}

including scheduled and unscheduled downtime---

@@

\subsection{ToO Events}
Describe the default ToO events we inject by default. xxx-was there a citeable thing that came out of the workshop?

@@

For the majority of observations we use a modified Markov Decision Process (MDP), similar to the one initially developed in \citet{Naghib2019}.

@@

Most surveys define a fixed set of fields (e.g., DES). Early attempts at scheduling LSST used fixed fields, but this leads to an issue of how to optimally dither pointings to fill chip and raft gaps, as well as distribute field overlap regions.

Rather than use a fixed set of specific pointings, we adopt a full-sky tesselation and then randomize the orientation of that tesselation nightly. So on any given night, there are XXX potential coordinates the telescope could be pointed to, but those positions will not be reused on subsequent nights.

@@

@@

There are concerns that LEO satellite mega-constellations could impact ground based astronomical observatories. We have not included any satellite dodging as earlier work found that even for an Gen 2 Starlink constellation with 30,000 satellites, only 0.04\% of LSST science pixels would be lost to satellite streaks \citep{Hu2022} (xxx-tech note too).

\software{Astropy \citep{astropy2013, astropy2018}, \\
          Astroplan \citep{astroplan18},
          Healpy and HEALpix\footnote{\url{http://healpix.sourceforge.net}} \citep{healpix2005, Zonca2019},
          Numpy \citep{harris2020array}, \\
          Matplotlib \citep{Hunter:2007},
          Scipy \citep{2020SciPy-NMeth},
          Pandas \citep{reback2022}, \\
          Scikit-learn \citep{scikit-learn},
          OpenOrb \citep{Granvik2009}, 
          GNU Parallel \citep{Tange2023}, \\
          rubin\_scheduler \citep{rubin_scheduler2024},
          rubin\_sim \citep{rubin_sim2024}}
